Symbolic Weighted Language Models and\\ Quantitative Parsing over Infinite Alphabets

### 2021-07-08 introduction.tex
Intro revisee Ph.

The latter case is related to a study that motivated the present work:
automated music transcription. Most representations of music are
essentially linear. This is true for audio files,
but also for widely used symbolic representations like MIDI.
Such representations ignore the hierarchical structures that frame the
conception of music, at least in the western area. These structures, on the other hand,
are present, either explicitly  or implicitly,
in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
in beats, and so on so forth. Music events do not occur at arbitrary timestamps,
but respect a discrete division of the  timeline incurred by
these recursive divisions.
 The \emph{transcription problem} takes
a linear input (audio or MIDI) and aims at re-constructing these structures
by mapping input events to this hierarchical rhythmic space.
It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
over an infinite alphabet of timed events.

>>>>

The latter case is related to a study that motivated the present work:
automated music transcription. Most representations of music are
essentially linear. This is true for audio files,
but also for widely used symbolic representations like MIDI.
Such representations ignore the hierarchical structures that frame the
conception of music, at least in the western area. These structures, on the other hand,
are present, either explicitly  or implicitly,
in music notation~\cite{Gould11Notation}: music scores are partitioned in measures, measures
in beats, and beats can be further recursively divided.
It follows that music events do not occur at arbitrary timestamps,
but respect a discrete division of the  timeline incurred by
these recursive divisions.
The \emph{transcription problem} takes
as input a linear repreentation (audio or MIDI) and aims at re-constructing
these structures
by mapping input events to this hierarchical rhythmic space.
It can therefore be stated as a parsing problem~\cite{foscarin:hal-01988990},
over an infinite alphabet of timed events. Throughout the paper we will
illustrate our framework with examples taken for this motivating application.

### 2021-07-08 semiring.tex
Intro revisee Ph.

% !TEX root = main.tex
%
% Semiring basics
%
We shall consider semirings for the weight values of our language models.
\florent{The results are established for a general class of semirings. They can be instantiated for concrete cases}
%
\philippe{There is sometimes a confusion in the text between the struture and the domain $\Semiring$. Not
essential}.
A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$ 
is a structure with a domain~$\Semiring$,
equipped with two associative
binary operators~$\oplus$ and $\otimes$,
with respective neutral elements $\zero$ and $\one$, and such that:
%$\< \mathbb{S}, \oplus, \zero>$ is a commutative monoid
%$\< \mathbb{S}, \otimes, \one>$ is a monoid
\begin{itemize}
\item $\oplus$ is commutative: 
 $\< \Semiring, \oplus, \zero>$ is a commutative monoid 
   and $\< \Semiring, \otimes, \one>$ a monoid,
\item $\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
$x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$, 
and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$,
\item $\zero$ is absorbing for~$\otimes$: 
$\forall x\in \mathbb{S}$, $\zero \otimes x = x \otimes \zero = \zero$.
\end{itemize}
%Components of a semiring~$\Semiring$ may be subscripted by~$\Semiring$ when needed.
%We simply write $x \in \Semiring$ to mean $x \in \mathbb{S}$.
%
Intuitively, in the models presented in this paper, 
$\oplus$ selects an optimal value from two given values, 
in order to handle non-determinism, 
and $\otimes$ combines two values into a single value, 
in a chaining of transitions.
%and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},

@@

In~\cite{Huang08advanceddynamic}, when the property~$(i)$ holds,  
$\Semiring$ is called \emph{superior} \wrt the ordering~$\leq_\oplus$.
We have seen in the proof of Lemma~\ref{lem:bounded} that it implies that 
$\one \leq_\oplus x$ for all $x \in \Semiring$.
Similarly, by the first inequality of ($i$) with $y = \zero$,  
$x \leq_\oplus x \otimes \zero = \zero$.
%
Hence, in a superior semiring, % \wrt~$\leq$, 
it holds that %$\one \leq \zero$
for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
%
Intuitively, from an optimization point of view,
it means that $\one$ is the best value, and $\zero$ the worst.
%** superior implies $\Semiring$ bounded~\cite{Mohri02semiring} see **
%
In~\cite{Mohri02semiring}, 
$\Semiring$ with the property ($ii$) of Lemma~\ref{lem:bounded}  
is called \emph{bounded} -- we shall use this term in the rest of the paper. 
% \emph{negative boundedness} 
It implies that, when looking for a best path in a graph whose edges
are weighted by values of $\Semiring$, the loops can be safely avoided,
because, for all $x \in \Semiring$ and $n \geq 1$, 
 $x \oplus x^n = x \otimes (\one \oplus x^{n-1}) = x$.

>>>>

% !TEX root = main.tex
%
% Semiring basics
%
We shall consider semirings for the weight values of our language models.
%\florent{The results are established for a general class of semirings. They can be instantiated for concrete cases}
A \emph{semiring} $\< \Semiring, \oplus, \zero, \otimes, \one>$
is a structure with a domain~$\Semiring$,
equipped with two associative
binary operators~$\oplus$ and $\otimes$,
with respective neutral elements $\zero$ and $\one$, and such that:
%$\< \mathbb{S}, \oplus, \zero>$ is a commutative monoid
%$\< \mathbb{S}, \otimes, \one>$ is a monoid
\begin{itemize}
\item $\oplus$ is commutative:
 $\< \Semiring, \oplus, \zero>$ is a commutative monoid
   and $\< \Semiring, \otimes, \one>$ a monoid,
\item $\otimes$ distributes over~$\oplus$:  $\forall x, y, z \in \mathbb{S}$,
$x \otimes (y \oplus z) = (x \otimes y) \oplus (x \otimes z)$,
and $(x \oplus y) \otimes z = (x \otimes z) \oplus (y \otimes z)$,
\item $\zero$ is absorbing for~$\otimes$:
$\forall x\in \mathbb{S}$, $\zero \otimes x = x \otimes \zero = \zero$.
\end{itemize}
%Components of a semiring~$\Semiring$ may be subscripted by~$\Semiring$ when needed.
%We simply write $x \in \Semiring$ to mean $x \in \mathbb{S}$.
%
Intuitively, in the models presented in this paper,
$\oplus$ selects an optimal value from two given values,
in order to handle non-determinism,
and $\otimes$ combines two values into a single value.
%and let $\< \Semiring, \oplus, \zero, \otimes, \one>$ be a {semiring},

@@

In~\cite{Huang08advanceddynamic}, when the property~$(i)$ holds,
$\Semiring$ is called \emph{superior} \wrt the ordering~$\leq_\oplus$.
We have seen in the proof of Lemma~\ref{lem:bounded} that it implies that
$\one \leq_\oplus x$ for all $x \in \Semiring$.
Similarly, by the first inequality of ($i$) with $y = \zero$,
$x \leq_\oplus x \otimes \zero = \zero$.
%
Hence, in a superior semiring, % \wrt~$\leq$,
it holds that %$\one \leq \zero$
for all $x \in \Semiring$, $\one \leq_\oplus x \leq_\oplus \zero$.
%
Intuitively, from an optimization point of view,
it means that $\one$ is the best value, and $\zero$ the worst.
%** superior implies $\Semiring$ bounded~\cite{Mohri02semiring} see **
%
In~\cite{Mohri02semiring},
$\Semiring$ with the property ($ii$) of Lemma~\ref{lem:bounded}
is called \emph{bounded} -- we shall use this term in the rest of the paper.
% \emph{negative boundedness}
\philippe{Ca j'ai pas compris}
It implies that, when looking for a best path in a graph whose edges
are weighted by values of $\Semiring$, the loops can be safely avoided,
because, for all $x \in \Semiring$ and $n \geq 1$,
 $x \oplus x^n = x \otimes (\one \oplus x^{n-1}) = x$.
